---
author: "Lucía García-Duarte Sáenz"
title: "Annex I"
output: html_document
---

*This notebook contains a deeper visual analysis of the NYC taxi trips during Thanksgiving 2018.*


## Spatial analysis (cont.)

### Chunk 1
Here it is shown how to extract the latitude and longitud values of each zone from the `.shp` file, so to obtain a complete look-up table containing the zone ID, zone name, borough and coordinates. Also, the weights of the links connecting each node (location) are obtained, which represents the total number of times a trip from one location to another has been done.

Data was saved and exported to represent the network of locations and most common trips using Gephi. Also, it has used in the main Python notebook for further analysis.

```{r node_link, message=FALSE, warning=FALSE}
df <- read.csv('data/data_filtered.csv',row.names=NULL)
library(dplyr)
library(plotly)
library(raster)
shp <- shapefile("data/tlc_yellow_geom.shp")# readOGR("data/tlc_yellow_geom.shp")
shp@data$seq_id <- seq(1:nrow(shp@data))
# Correct mistake (it can be seen in the file 'taxi+zonelookup.csv' that these three are the same)
c1 = which(as.integer(shp@data$zone_id)==103)
shp@data$zone_id[c1[2]] = '104'
shp@data$zone_id[c1[3]] = '105'

# in order to plot polygons, first fortify the data
shp@data$id <- rownames(shp@data)
# create a data.frame from our spatial object
shpdata <- fortify(shp, region = "id")
# merge the "fortified" data with the data from our spatial object
shpdf <- merge(shpdata, shp@data, by = "id")


### NODE
nodes = data.frame(shpdf$zone_id,shpdf$zone_name,shpdf$borough,shpdf$long,shpdf$lat)
colnames(nodes) <- c('Id', 'Name', 'borough', 'longitude','latitude')
nodes <- nodes %>% group_by(Id,Name,borough) %>% summarise('longitude' = mean(longitude),'latitude' = mean(latitude))
write.csv(nodes,'data/nodes.csv',row.names=FALSE)

### LINK
trips = as.data.frame(paste(df$pickup_location_id, df$dropoff_location_id, sep="_")) 
colnames(trips) <- 'value' 
#Count path combinations 
weights=trips %>% 
  group_by(value) %>% 
  dplyr::summarise(weight=n())
## Define links or edges (source, target, weight)
links <- data.frame(do.call(rbind, strsplit(as.vector(weights$value), split = "_"))) #Separate start and end station IDs
names(links) <- c("source", "target")
links$weight<-weights$weight 
write.csv(links,'data/links.csv',row.names=FALSE)
```

### Chunk 2

Different maps showing specific features per location.
1. Amount of pick-ups 
2. Amount of drop-offs
3. Most common picked-up number of passengers
4. Most common dropped-off number of passengers

```{r message=FALSE, warning=FALSE}
### Picks and drops count ###
picks_count <- df %>% group_by(pickup_location_id) %>% dplyr::summarise(ntrips_pu = n())
drops_count <- df %>% group_by(dropoff_location_id) %>% dplyr::summarise(ntrips_do = n())

aux=as.data.frame(as.integer(shp@data$zone_id))
colnames(aux) <- 'pickup_location_id'
picks_count = aux %>% left_join(picks_count)
colnames(aux) <- 'dropoff_location_id'
drops_count = aux %>% left_join(drops_count)

shp@data$picks_count <- picks_count$ntrips_pu
shp@data$drops_count <- drops_count$ntrips_do

### Most common number of passengers ###
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
picks_pass = df %>% group_by(pickup_location_id) %>% dplyr::summarise(npass_pu = getmode(passenger_count))
drops_pass = df %>% group_by(dropoff_location_id) %>% dplyr::summarise(npass_do = getmode(passenger_count))

colnames(aux) <- 'pickup_location_id'
picks_pass = aux %>% left_join(picks_pass)
colnames(aux) <- 'dropoff_location_id'
drops_pass = aux %>% left_join(drops_pass)

shp@data$picks_pass <- as.character(picks_pass$npass_pu)
shp@data$drops_pass <- as.character(drops_pass$npass_do)


# in order to plot polygons, first fortify the data
shp@data$id <- rownames(shp@data)
# create a data.frame from our spatial object
shpdata <- fortify(shp, region = "id")
# merge the "fortified" data with the data from our spatial object
shpdf <- merge(shpdata, shp@data, by = "id")

```

```{r message=FALSE, warning=FALSE}
p0 <- ggplot() +
geom_polygon(data = shpdf, aes(x = long, y = lat, group = group, fill = picks_count, text = paste("Zone:",zone_name,"\nBorough:",borough,"\nCount:",picks_count)), color="white", size = 0.2) +
scale_fill_distiller(palette = "Greens", name = 'Count') +
coord_equal(ratio = 1.3) +
theme(panel.background = element_rect(fill = '#D5EEFF', color = 'purple'), plot.title = element_text(vjust = -10)) +
labs(title = "Amount of pick-ups per zone in NYC during Thanksgiving 2018", x = "Longitude", y = "Latitude")
p0 %>% ggplotly(tooltip = "text")

p1 <- ggplot() +
geom_polygon(data = shpdf, aes(x = long, y = lat, group = group, fill = drops_count, text = paste("Zone:",zone_name,"\nBorough:",borough,"\nCount:",drops_count)), color="white", size = 0.2) +
scale_fill_distiller(palette = "Greens", name = "Count") +
coord_equal(ratio = 1.3) +
theme(panel.background = element_rect(fill = '#D5EEFF', color = 'purple'), plot.title = element_text(vjust = -10)) +
labs(title = "Amount of drop-offs per zone in NYC during Thanksgiving 2018", x = "Longitude", y = "Latitude")
p1 %>% ggplotly(tooltip = "text")
```

```{r message=FALSE, warning=FALSE}
p0 <- ggplot() +
geom_polygon(data = shpdf, aes(x = long, y = lat, group = group, fill = picks_pass, text = paste("Zone:",zone_name,"\nBorough:",borough,"\nMode:",picks_pass)), color="white", size = 0.2) +
coord_equal(ratio = 1.3) +
theme(panel.background = element_rect(fill = '#D5EEFF', color = 'purple')) +
labs(title = "Most common picked-up number of passengers per zone", x = "Longitude", y = "Latitude", fill = "Mode")
p0 %>% ggplotly(tooltip = "text")

p1 <- ggplot() +
geom_polygon(data = shpdf, aes(x = long, y = lat, group = group, fill = drops_pass, text = paste("Zone:",zone_name,"\nBorough:",borough,"\nMode:",drops_pass)), color="white", size = 0.2) +
coord_equal(ratio = 1.3) +
theme(panel.background = element_rect(fill = '#D5EEFF', color = 'purple')) +
labs(title = "Most common dropped-off number of passengers per zone", x = "Longitude", y = "Latitude", fill = "Mode")
p1 %>% ggplotly(tooltip = "text")
```

